---
title: "07 Identifying Specific Charges"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{07 Identifying Specific Charges}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, echo = FALSE, message=FALSE}
library(ojodb)
```

Many of the questions we ask require us to focus on specific criminal charges (*e.g*., [drug possession](https://okpolicy.org/in-its-first-year-sq-780-reversed-10-years-of-growth-in-felony-filings/) or burglary) or civil issues (*e.g.*, [evictions](https://openjustice.okpolicy.org/blog/oklahoma-court-tracker/)). Because the data we collect is not always clean or entered uniformly, identifying cases requires some patience and diligence. This document provides a process to do that.

## Identifying criminal charges

Let's say we want to look at cases with a top charge of second degree burglary that were filed in 2018. We'll start by querying all of the criminal felony cases for that year:

```{r}
connect_ojo()

d <- ojo_tbl("ojo_crim_cases") %>% 
  filter(casetype == "CF",
         file_year == 2018) %>% 
  collect()

disconnect_ojo()
```

This gives us 36,561 cases. The ojo_crim_cases table is a summary of all the charges and dispositions in a case. A criminal case is defined as all of the charges filed against a specific *person* in a specific court on a specific date, so each case has a unique combination of the variables `court`, `defname`, and `file_date`.

The `top_ct_desc` variable contains a description of the top count associated with the case. Ideally, we would have a systematic way to standardize charges from their dirty, hand-entered form; connect them to their statutory definitions; and rank them in terms of severity by finding the maximum sentence for each charge. But we have not developed that system yet, so the `top_ct_desc` is simply the first count description that appears on the case record.

We'll use the `top_ct_desc` variable to find the second degree burglary cases we're interested in.

### Start broad, then narrow down

To identify burglary charges, we'll use the `str_detect()` function on the `top_ct_desc` variable to look for certain words or parts of words and exclude other words using `!str_detect()`. This code creates a new variable, `burg2`, which tells us whether or not the `top_ct_desc` variable contains the string "BURGLARY".

```{r}
d <- d %>% 
  mutate(burg2 = str_detect(top_ct_desc, "BURGLARY"))
```

We can then look at all of the charges that were caught as burglary and count the number of occurrences like this:

```{r}
d %>% 
  filter(burg2 == TRUE) %>% # Look just at the charges that fit our criteria
  count(top_ct_desc) %>%    # Count the frequency of each unique value of top_ct_desc
  arrange(desc(n))          # Sort them in descending order of frequency (n)
```

There are 211 different ways that top charges of burglary were recorded in 2018, which gives a sense of how difficult this data is to standardize. To look at the full list, you can add `%>% view()` at the end of code above.

Since we simply searched for "BURGLARY", the `burg2` variable does not tell us whether the charge is first, second, or third degree. We see from the above list that "SECOND" and "2ND" both denote second degree burglary, so we'll narrow to just those charges by including another `str_detect()`:

```{r}
d <- d %>% 
  mutate(burg2 = str_detect(top_ct_desc, "BURGLARY") & str_detect(top_ct_desc, "SECOND|2ND"))
```

Then we can check our list again:

```{r}
counts <- d %>% 
  filter(burg2 == TRUE) %>% 
  count(top_ct_desc) %>% 
  arrange(desc(n))

sum(counts$n)
```

This gives us 2,297 second degree burglary cases. Looking through the charges that were classified as second degree burglary, it looks like this definition holds up. But since there are typos, misspellings, and abbreviations in charge descriptions sometimes, we want to try pieces of words as well:

```{r}
d <- d %>% 
  mutate(burg2 = str_detect(top_ct_desc, "BURG") & str_detect(top_ct_desc, "SECOND|2ND"))

counts <- d %>% 
  filter(burg2 == TRUE) %>% 
  count(top_ct_desc) %>% 
  arrange(desc(n))

sum(counts$n)
```

Shortening "BURGLARY" to "BURG" found us 5 more cases. That's not a lot, but it's good to keep in mind that bits of words can find charges that whole words will not.

## Identifying charges by exclusion

In some cases, it's not entirely feasible to anticipate all the possible ways an offense might be recorded. In particular, cases involving drugs may refer to the substances in general as "DRUGS", "CDS", "CONTROLLED SUBSTANCE", or specifically as "METH", "METHAMPHETAMINE", "MARIJUANA", "MARI", etc. Sometimes it's more effective to exclude charges that contain certain words rather than trying to identify all of the possible ways something might appear.

If we want to identify possession of a controlled substance with intent to distribute (PWID), we can start by looking for "INTENT" or "DISTR":

```{r}
d <- d %>% 
  mutate(pwid = str_detect(top_ct_desc,"INTENT|DISTR"))

counts <- d %>% 
  filter(pwid == TRUE) %>% 
  count(top_ct_desc) %>% 
  arrange(desc(n))
```

This simple search is effective; 19 of the top 20 on the list are what we're looking for. There are also some charges like "Shooting with intent to kill" and "Entering with intent to steal copper" that we want to exclude. Taking stock of those, we can add words to exclude, run it again, find new offenses we didn't catch, and repeat the process. We might gradually expand it to something like this:

```{r}
d <- d %>% 
  mutate(pwid = str_detect(top_ct_desc,"INTENT|DISTR") & 
                                   !str_detect(top_ct_desc, "MISDEMEANOR|HUMAN|VEHICLE|FIREARM|CULTIVATION|MURDER|PORN|STEAL|SHOOT|ASSAULT|BREAK|KILL|BATT"))

counts <- d %>% 
  filter(pwid == TRUE) %>% 
  count(top_ct_desc) %>% 
  arrange(desc(n))
```

This is an inexact process for now, and we need to be transparent about what we're doing. Since our database isn't open to the public, we'll include a csv file on the GitHub repository with a list of all of the raw charges and their frequency for each analysis that uses this classification method. To do that, write the `counts` dataframe produced above to a csv:

```{r, eval=FALSE}
write_csv(count, "PWID charge descriptions.csv")
```


