---
title: "Overview of the OJO database"
output: rmarkdown::html_vignette
description: >
  Learn what's in the OJO database and how it's organized
vignette: >
  %\VignetteIndexEntry{Overview of the OJO database}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE}
library(tibble)
library(htmlTable)
```

The Open Justice Oklahoma database contains a wide range of data related to the state's criminal and civil justice systems gathered from a variety of sources, as well as supplemental data generated by OJO processes, such as identity-linking tables and geocoded addresses of evictions. It is a MySQL database hosted on Amazon Web Services, and it is currently accessible only by OJO staff and authorized partners through a database connection. We hope to release an API to allow public access to the data at some point.

The goal of the ojo R package is to provide tools to access, clean, and analyze this data with standardized methods.

## Data sources

The name of each table in the OJO database begins with a prefix signifying the source of the data, as outlined below.

### Raw data

Most data in the OJO database is stored in the same form it was collected. Since it is administrative data, it is not standardized and includes a healthy amount of typos and other errors. We try to preserve the raw data from every source in the database and layer our cleaning and standardizing methods on top of the raw data. 

The table below describes the sources of raw data.

``` {r, echo = FALSE}
htmlTable::htmlTable(tibble(Prefix = c("oscn_",
                                       "odcr_",
                                       "doc_",
                                       "dlm_"),
                            `Data related to` = c("District Courts (13 counties, including 6 largest by population)",
                                                  "District Courts (all courts not included in OSCN)",
                                                  "Prisons",
                                                  "Tulsa County Jail"),
                            Source = c("http://www.oscn.net",
                                       "https://www1.odcr.com/",
                                       "http://doc.ok.gov/odoc-public-inmate-data",
                                       "http://iic.tulsacounty.org/expInmateBookings")),
                     align = "l")
```

### OJO-generated data

Tables with the prefix "ojo_" relate to all aspects of the justice system. These tables store data generated by OJO methods and are used to supplement raw data.

